---
title: "Installation"
author: Luís Valente
---

## Install required software:

-   [R](https://www.r-project.org) **Install latest version of R if possible**

-   [Rstudio](https://posit.co)

-   [Figtree](https://github.com/rambaut/figtree/releases)

-   [Geneious](https://manage.geneious.com/free-trial) **Free trial lasts for 30 days**

-   [Beast 2](https://www.beast2.org) (Includes Beast, Beauti, TreeAnnotator)

-   [Tracer](https://github.com/beast-dev/tracer/releases/tag/v1.7.2 )



## Install required R packages

Open and run this script in Rstudio. This should install all required R packages (**may take a while!**).

```{r install, eval=F}
install.packages('ape',dependencies = TRUE)
install.packages('phytools',dependencies = TRUE)
install.packages('paran')
install.packages('splits', repos='http://R-Forge.R-project.org')
```

If you have problems with installing package dependencies, this could be because you need to install a more recent version of R.

Check if packages load correctly:

```{r loadp, eval=F}
library(ape)
library(phytools)
library(splits)
```

