---
title: "Species delimitation"
author: Luís Valente
---

```{r, echo=F}
library(ape)
cyanistes<-read.nexus('Cyanistes.tre')
```

## Species delimitation using the GMYC model

How many species of blue tits are there? We will use a model to delimit species based on the divergence between clades in the phylogeny. The GMYC model (Fujisawa & Barraclough 2013, *Systematic Biology*), which stands for "Generalized Mixed Yule Coalescent Approach", is a likelihood method for delimiting species by fitting within- and between-species branching models to reconstructed gene trees.

Load the `splits` package ("SPecies' LImits by Threshold Statistics"), which includes an R implementation of the GMYC model.

```{r setup3, warning=FALSE,message=FALSE}
library(splits) 
```

In case you haven't done so above, remove the outgroup (delete all *Parus* individuals), by creating a new object without all the tips belonging to the outgroup.

```{r}
cyanistes_ingroup<-drop.tip(cyanistes,c("Parus_major_major_KP759174.1",
                                         "Parus_major_DQ792786.1", 
                                         "Parus_major_DQ792787.1",
                                         "Parus_major_EU167009.1", 
                                         "Parus_major_KJ456375.1"))
```

Now we can fit the GMYC model to the data.

```{r, results="hide"}
cyanistes_gmyc <- gmyc(cyanistes_ingroup)
```

Plot the statistics. How many clusters are identified? What do these clusters mean?

```{r, results="hide"}
summary(cyanistes_gmyc)
```

View which tips in the tree are assigned to each cluster

```{r, results="hide"}
spec.list(cyanistes_gmyc)
```

Visualize the results of the GMYC model (Hit return to view the different plots). **May require some adjusting on your screen.**

```{r, eval=F}
plot(cyanistes_gmyc)
```

-   How many "species" does the model identify?

-   Do the clusters match species based on the tip names?

-   Would you recommend a taxonomic revision of this group?

-   How many **endemic** species of *Cyanistes* are there in the Canary Islands?

-   How many **endemic subspecies**? 

-   What other sources of evidence would be useful to define species of blue tits in the Canarian archipelago?

-   What species would you give priority to conservation (if any)?
